What's hot and what's not:\\Windowing developer topic analysis

### 2009-04-10 lda-paper.tex
some ric edits plus new month plot

As development on a software project progresses, developers will naturally
shift their focus between different topics and tasks many times.  Managers
and newcomer developers often seek simple ways of understanding what tasks
have recently been worked on and how much effort has gone into each; for
example, a manager might wonder what unexpected tasks occupied their team's
attention during a period when they were supposed to have been implementing
a set of new features.  Tools such as Latent Dirichlet Allocation (LDA) and
Latent Semantic Indexing (LSI) can be used to analyze check-in comments
over the entire lifetime of a project, and previous work on developer topic
analysis has leveraged these tools to associate commit log comments with
independent topics extracted from these commit log comments.  In this work, we use
LDA and LSI to analyze periods, such as months, within a project's lifetime
to create a time-windowed model of changing development topics.  We propose
a visualization of this model that allows us to explore the evolving stream
of topics of development occurring over time, and we demonstrate that
windowed topic analysis offers strong advantages over entire lifetime topic
analysis because of the independence of topics between windows.

@@

Topic analysis, with respect to Software Control Systems (SCSs) such
as CVS, Subversion, Git etc., is potentially valuable as it has many
uses for various stakeholders and uses.  Imagine a development team
has agreed to implement some features at the start of an iteration. By
the end of the iteration only some of the agreed upon features are
implemented. The developers insist they were focusing solely on the
features agreed upon, but what actually happened? Topic analysis could
help by teasing out the distinct topics or focusses of the
developers. Perhaps the developers had to deal with issues.
Developers might have grouped those orthogonal changes into one of the
features they were working on, thus when asked about how their time
was spent the orthogonal changes would not be covered. This is the kind of underlying story that we
hope that windowed topic analysis can uncover. If windowed topic
analysis can discover important but overlooked topics, it can help
both developers and managers. Developer can see a summary of what
worked on, and managers can development.

Commits in SCSs are not necessarily consistent, they can have multiple
purposes, thus we hope that these multiple purposes manifest themselves as
topics interleaved within the commits.   Others have demonstrated that by
using powerful topic analysis techniques, such as Latent Semantic Indexing
(LSI) and Latent Dirichlet Allocation (LDA), can aid in separating topics
from these mixed comments. These topics often describe distinct ideas
that are interleaved throughout the development. A topic represents a
groups of commit comments that are related to each other by their
content.  In this case a topic is a set of tokens.

@@

One use of topic analysis is to help us partition a project's
time-line into periods of developer focus. By breaking apart an
iteration into sets of topics and trends (topics which recur), we may
be able to recognize the underlying software development process from
these commits. Alternatively, we can determine what particular
maintenance task was being performed are a given time.

We propose to extend the use of common topic analysis techniques by
executing them over windows of documents over time. We hope by looking
at topic clusters during periods or windows over time we can identify
continuous effort on a select group of topics as well reflect topics
that were local to that period. Figure \ref{fig:lda} is an example of
the kind of visualization we want to automatically generate. We want
to see topics that are unique to a certain time period, displayed
across the time axis. Those topics that recur, or occur over a
larger period are plotted continuously. In our example we have
summarized each topic with a single word.

In this paper we explore how topics shift over time in the source
control repository of a project, using several open source database
systems as examples. We analyze commit comments per revision in each
time window and we extract the topics in that time window. We expect
that topics in each window will change over time and the similarities
and differences in these sets of topics will indicate both developer
focus and changes in developer focus.

@@

We now define some terms that we will use in the rest of the paper: A
\emph{message} is a block of text written by developers. In this
paper, messages will be the CVS commit comments made when the user
commits changes to files in a repository. A \emph{word distribution}
is the summary of a message by word count. The word distribution will
be over all messages looked at but since most words will not appear in
a message, a word distribution is effectively a word count divided by
the message size. A \emph{topic} is essentially a set of words that
form a word distribution that is unique and independent within the set
of documents in our total corpus. One could think of a topic as the
center distribution of a group of messages. In this paper we often
summarize topics by their top 10 most frequent words in their word
distribution.  A \emph{trend} is a one or more similar topics which
recur over time.  Trends are particularly important, as they indicate
long-lived and recurring topics that may provide key insights into the
development of the project.

@@

In terms of clustering and finding topic distributions, Latent
Dirichlet Allocation (LDA)~\cite{944937} competes with Latent Semantic
Indexing
(LSI)~\cite{1421013,1374321,10.1109/ICPC.2007.13,10.1109/ICPC.2006.17},
probabilistic Latent Semantic Indexing (pLSI)~\cite{944937} and
semantic clustering~\cite{1698774,1566153}. These tools are used for
document modelling, document clustering and collaborative
filtering. What LDA attempts to do is encode documents as mixture
model of topics.  LDA is used in software engineering
literature~\cite{lukins2008,10.1109/MSR.2007.20,NIPS2007637,1321709}
is to extract topic clusters from documents such as methods, bug
reports, source code and files.

@@

LSI is related to LDA and has been used to identify topics in software
artifacts for formal concept analysis and concept
location~\cite{1421013,1374321,10.1109/ICPC.2007.13,10.1109/ICPC.2006.17}.
Concept analysis aims to automatically extract concepts from source
code or documents that are related to the source code.  Concept location concerns how
high level concepts relate to low level entities such as source code, for
example, when fixing a bug how and where do the concepts in the bug
report map back to the source code?  Semantic Clustering has also been
used for similar purposes~\cite{1698774,1566153} as it is similar to
LSI.

@@

Our technique differs from the previous techniques because we apply
LDA locally to month-long windows of commit comments, where as other
approaches apply LDA once to the entire project. This allows us to
examine the time-based windowing of topics over its development
history.

@@

In our first exploratory pass we wanted to see if LDA could provide
interesting topics extracted from a real system. We took the
repository of MySQL 3.23, extracted the commits, grouped them by 30
day non-overlapping windows, and then applied LDA to each of these
windows. We asked LDA to make 20 topics per window and we then
carefully examined the top 10 most frequent words in that topic.  We
found there were common words across topic clusters, such as
\emph{diffs}, \emph{annotate} and \emph{history}, these words probably should
be viewed as stop words. There were notable transitional topics such
as in the first window the word \emph{bitkeeper} appears because MySQL
adopted Bitkeeper for their source control system yet in the following
windows there were no mentions of Bitkeeper. \emph{RENAME} also only
appeared once in the first window and never again. Per each topic we
tried to identify the purpose of the topic; to our surprise we found
that even with only a little familiarity with the code base that
naming the topic was straightforward.

A sub-sampling of the notable topic words is displayed in Table
\ref{tab:portability}.  To name the topics we selected terms that
seemed to best summarize those topics.  After viewing these topics, it was
obvious that we would have to combine try to track the evolution of
topics over the windows or rank topics by similarity, thus if they
were similar enough we joined them into a trend. Figure \ref{fig:lda}
displays a sample plot of the extracted topics in table
\ref{tab:portability}.

@@

2001 &  Aug &      TABLES row version \\
\hline         
2001 &  Sep &      update checksum merge \\
% 2001 &  Oct &      fixed fix \\
% 2001 &  Dec &      HPUX SCO fix \\
% \hline         
% 2002 &  Feb &      net buffer length  max\_allowed\_packet \\
% 2002 &  Mar &      small buf fix  \\
% \hline         
% 2002 &  May &      [popular] fix SCO OSF1 table\_name \\
% 2002 &  Nov &      HPUX11 compiler HP \\
% \hline         
% 2003 &  Feb &      Linux errno   \\
% 2003 &  Mar &      alarm bookmark bug \\
% \hline         
% 2003 &  Sep &      Auto logging merge windows distribution fix 64-bit 4.0 Cleanup \\
\hline
\end{tabular}
\caption{Sub-sampled topics from MySQL 3.23, some with continuous topics, these tokens were pulled from LDA topic. Each token is a summary of one LDA generated topic from MySQL 3.23 commit logs.}
\label{tab:portability}
\end{table}

\begin{figure*}
  \centering
  \includegraphics[width=0.9\textwidth]{lda}
  \caption{Example of topics extracted from MySQL 3.23. This is the kind of plot we eventually want to produce: named topics and topic trends. The horizontal axis is time in months and the vertical axis used to stack topics that occur at the same time. Longer topics are topics which recur in adjacent window. Colors are arbitrary.}
  \label{fig:lda}
\end{figure*}

\section{Methodology}
\begin{figure*}
  \centering
  \includegraphics[width=0.9\textwidth]{commit-to-topics} 
  \caption{How commits are analyzed and aggregated into Topics and Trends. Commits are first extracted, then abstracted into word counts or word distributions. These word distributions are then given to topic clusterer like LDA which finds independent word distributions (topics) that these documents are related too and clusters or associates these documents with these topics.}
  \label{fig:commits}
\end{figure*}

@@

\subsection{Extract Repositories}

@@

%   2. [ ] Extract Documents
\subsection{Extract Documents}
%   3. [ ] Windowed Sets

@@

Given all messages, we group the messages into windows. We could use
overlapping windows, but in this paper we use non-overlapping windows
of a set time unit because it simplifies analysis.  Windowing by time
allows for many levels of granularity. We usually used one month as
the length of our time windows, while we could use different
granularities for the study we think that a month is a sizable unit of
development, which we consider to be fine enough to show meaningful
shifts in focus, but coarse enough not to show too much detail.

@@

Once we have our data windowed, we apply our Topic Analysis tool to
each window and extract the top $N$ topic clusters, we used $20$
topics in our case studies. Our Topic Analysis tool is an
implementation of LDA, but we could have used other similar tools like
LSI, but previous work showed more promising topic analysis results
with LDA.  We note that this step is a slow one, as executing even one
instance of LDA involves significant computation, and we perform LDA
once per window.

@@

%   6. [ ] Cluster Similarity [0/1]
\subsection{Cluster Similarity}
%      1. [ ] need the tool
%             grab from the cluster analysis
%   7. [ ] Continuous Cluster Analysis

Once we have our topics extracted for each window, we analyze them and
look for repeating topics. We find repeating topics by comparing the
10 most common words of each topic.

We then abstract these topics by comparing them to each other.  Each
topic is compared to each other topic in the system, given a certain
threshold of top 10 similarity, such as 8 out of 10 words. Then given
this topic similarity matrix, we find the transitive closure of
similar topics; that is, if we model topics are nodes and arcs as
similarity, we fill flood along the similarity arcs until we have
partitioned the topics into clusters of similar topics. Figure
\ref{fig:closure} provides an example of clustering topics by the
topic similarity. These clusters of topics are called trends. A trend
is probably interesting and valuable if there is more than one topic
in the trend.

@@

We have devised several ways to visualize these topics and trends.
For all of these plots if we find trends that have continuous
segments, then we plot those segments as joined across time.  We have
a compressed stacked trend view, shown in Figure \ref{fig:topicsmear}
and Figure \ref{fig:zoomedsmear}, which displays trends as they occur
on the time-based x-axis (placement along the y-axis is arbitrary).
Another plot we have is the trend-time line plot, shown in Figure
\ref{fig:trendtimeline}, each trend gets it own distinct y-value ,
while the x-axis is time; these topics are then each plotted on their
own line across time as they occur. Our final plot is the histogram
plot, shown in Figure \ref{fig:histogram} where we plot each trend on
its own line but stack up the segments of the trend, much like a
histogram.

The first visualization is a compressed view
(Figure\ref{fig:topicsmear} and Figure \ref{fig:zoomedsmear}) where we
try to sink the larger trends to the bottom.  Once the larger trends
are stacked we fill in the gaps with the smaller trends in the same
window, and then stack the smaller trends on top.  Although there is a
chance that gaps will be left due to non-optimal stacking in practice
there are lot of small trends (90\% of all trends contain 1 topic)
which fill in these gaps quickly.  Different instances of the same
trend share the same color; apart from that, the color of trends is
randomly chosen and color similarity is not meaningful.  The
compressed view is convenient because it shows all of the information
and the repeating continuous trends are easy to pick out, although
repeating discontinuous trends are harder to spot.

The second visualization is the trend-time line plot (Figure
\ref{fig:trendtimeline}), it displays repeating trends more clearly by
dedicating a horizontal line for trend segments belonging to one
trend. This means that if a trend contains discontinuous segments then
the segments appear on the same line.  However, the least common
trends need to be pruned away or the view will be very long.

The third visualization is the histogram plot (Figure
\ref{fig:histogram}) which superficially resembles the time-trend
plot.  However, in this view the trends are plotted together by
stacking to the left of their row, thus time information is lost.  The
trends are ordered by the number of topics in the trend.  The
histogram plot shows the count of instances of a trend and thus
indicates which trends occur the most. Unfortunately due to the large
number of topics given allotted space it is often best to crop off the
trends with only one topic, otherwise the tail is long.

%XXX Describe something useful like a trend
\begin{figure*}
  \centering
  \includegraphics[width=1.0\textwidth]{fixed-time-smear-plot-scaled}
  \caption{Compact trend plot shows topics per month for MaxDB
    7.500. The x-axis is time in months, the Y axis is used to stack
    topics occurring at the same time. Trends that are continuous are
    plotted as continuous blocks. The top 10 words in the topics are
    joined and embedded in box representing the topics.}
  \label{fig:topicsmear}
\end{figure*}

\begin{figure}
  \centering
  \includegraphics[width=0.45\textwidth]{fixed-time-smear-plot-cropped}
  \caption{A zoomed in slice of compact trend plot of topics per month of MaxDB 7.500. The topic text is visible in each topic box. Trends are plotted across time continuously.}
  \label{fig:zoomedsmear}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=0.45\textwidth]{class-smear-plot-crop-scaled}
  \caption{Trend-time Line: Trends plotted per month of MaxDB 7.500. Time in months are plotted along the x-Axis, each row on the y-axis is associated with a trend ranked by size in descending order}         
  \label{fig:trendtimeline}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=0.45\textwidth]{histogram-cropped-scaled}
  \caption{The top part of a trend histogram of MaxDB 7.500, ordered
    by topic occurrence. X axis determines the number of continuous
    months of a trend. Trends are ranked by the number of topics that
    a trend contains in descending order.}
  \label{fig:histogram}
\end{figure}

@@

The largest common trend has references to build system files like
\texttt{SYSDD.punix} and \texttt{MONITOR.punix}.  This trend is
partially displayed at the top of the zoomed in compressed view plot
(Figure \ref{fig:zoomedsmear}) and at the top of the histogram plot
(Figure \ref{fig:histogram}).  Other tokens mentioned are
\emph{Sutcheck v1.09}, the prefix SUT stands for Storable Unit
Type. Sutcheck would also automate check-ins using a Perforce SCS tool
which was exporting check-ins to CVS.

@@

We noticed that Perforce and Changelist were in just about every
topic, so we added them to the stop words list. Longer trends were
shortened and in some cases the total number of topics found per month
was less. Evaluating different similarity thresholds showed that by
removing common words one reduces the general similarity of topics,
that said the larger topics still existed. Thus if you add more
relevant stop words one should tune your topic similarity parameters
to handle these changes.

@@

\subsection{Compare to topics over entire development}

@@

We attempted the same kind of analysis, as shown in
Figure~\ref{fig:statictopics}. We extracted 20 topics and then plotted
the number of messages per month that were related to that topic. We
found was that one topic would often dominate the results, while
messages related to other topics did not appear very often.
% This technique seems useful if the 20 extracted topics are relevant
% and useful, that said one weakness of this approach is that documents
% for a topic might only occur in one time window, yet we are tracking
% messages that could be related to it over the entire lifetime of the
% project.
This approach seems reasonable if most of the extracted topics are of
broad interest during most of the development process.  However, it
may be that some topics are of strong interest but only briefly; in
such a case, a window-based model gives a much stronger indication of
the fleeting importance of such topics, and can help to put such a
topic into its proper context.

@@

If we approach the difference of static global topic analysis and
windowed topic analysis via common tokens we can see that common token
tend to dominate both results, at least in the case of MaxDB 7.500 our
local topic approach netted us topics which contained these relatively
important and common words: \emph{UNICODE}, \emph{DEC/OSF1},
\emph{ODBC}, \emph{crash}, \emph{SQLCLI}, \emph{SYSDD.cpnix},
\emph{backup}, \emph{select}, \emph{make}, \emph{memory}, \emph{view},
and finally \emph{debug}. None of these tokens were part of the static
topic analysis topics, but they were part of 566 commits (6\% of the
entire system) to MaxDB 7.500. These tokens were part of 87 out of 500
(25 months, 20 topics per month) of our locally generated topics.

@@

\begin{figure}
  \centering
  \includegraphics[width=0.45\textwidth]{maxdb7500-everything-by-month}
  \caption{MaxDB 7.500 Topics Analyzed with Static Topic Analysis, 20
    topics over the entire development history of MaxDB 7.500. The
    color of the topic indicates the number of documents matching that
    topic in that month relative to the number of documents (white is
    most, black is least).}
  \label{fig:statictopics}
\end{figure}

@@

We presented multiple visualizations that focused on different aspects
of the data: temporality of trends, trend size, and a compressed
overall trend view. The compressed trend view shows much more
information than the views that global analysis could show, it
indicates how focused a period is by the total number of topics as
well shows topics by similarity so one can track trends across time.
Our histogram trend-view immediately hi-lights and measures the size of
trends while our time-line trend view shows how a topic reoccurs over
time. These visualizations help us get an over feeling about the
common and unique topics that developers focus on. If implemented
properly a user can easily zoom and query a topic.

>>>>

  As development on a software project progresses, developers will
  naturally shift their focus between different topics and tasks many
  times.  Managers and newcomer developers often seek ways of
  understanding what tasks have recently been worked on and how much
  effort has gone into each; for example, a manager might wonder what
  unexpected tasks occupied their team's attention during a period
  when they were supposed to have been implementing a set of new
  features.  Tools such as Latent Dirichlet Allocation (LDA) and
  Latent Semantic Indexing (LSI) can be used to analyze check-in
  comments over the entire lifetime of a project, and previous work on
  developer topic analysis has leveraged these tools to associate
  commit log comments with independent topics extracted from these
  commit log comments.  In this work, we use LDA and LSI to analyze
  periods, such as months, within a project's lifetime to create a
  time-windowed model of changing development topics.  We propose a
  visualization of this model that allows us to explore the evolving
  stream of topics of development occurring over time.  We demonstrate
  that windowed topic analysis offers advantages over topic analysis
  applied to a projects lifetime because many topics are quite local.

@@

Topic analysis, with respect to Software Control Systems (SCSs) such
as CVS, Subversion, Git etc., is potentially valuable as it has many
uses for various stakeholders and uses.  Imagine a development team
has agreed to implement some features at the start of a development
iteration. By the end of the iteration only some of the agreed upon
features are implemented. The developers insist they were focusing
 on the features agreed upon, but what actually happened? Topic
analysis could help by teasing out the distinct topics or focusses of
the developers. Perhaps the developers had to deal with other pressing issues.
Developers might have grouped those orthogonal issues into the
features they were working on, thus when asked about how their time
was spent the orthogonal changes would not be covered. This is the
kind of underlying story that we hope that windowed topic analysis can
uncover. If windowed topic analysis can discover important but
overlooked topics, it can help both developers and managers. Developers
can see a summary of what worked on, and managers can development.

Commits in SCSs are not necessarily consistent and they can have
multiple purposes. Thus we hope that our research will help uncover
these multiple purposes that manifest themselves as topics interleaved
within the commits.  Others have demonstrated that by using 
topic analysis techniques, such as Latent Semantic Indexing (LSI) and
Latent Dirichlet Allocation (LDA), can aid in separating topics from
these mixed comments. These topics often describe distinct ideas that
are interleaved throughout the development. A topic represents a
groups of commit comments that are related to each other by their
content.  In this case a topic is a set of tokens.

@@

One use of topic analysis is to help us partition a project's
time-line into periods of developer focus. By breaking apart an
iteration into sets of topics and trends (topics which recur), we may
be able to recognize the underlying software development process from
these commits. Alternatively, we can determine what particular
maintenance task was being performed are at a given time.

We propose to extend the use of common existing topic analysis techniques by
applying them to windows of documents over time. We hope by looking
at topic clusters during periods or windows over time we can identify
ongoing areas of focus on a select group of topics as well reflect more intermittent topics. Figure \ref{fig:lda} illustrates
the kind of visualization we want to automatically generate. We want
to see topics that are unique to a certain time period, displayed
across the time axis. Those topics that recur, or occur over a
larger period are plotted continuously. In our example we have
titled each topic with a single word.

In this paper we explore how topics shift over time in the source
control repository of a project, using several open source database
systems as examples. We analyze commit comments per revision in each
time window and we extract the topics in that time window. We expect
that topics will change over time and the similarities
and differences in these topics will indicate developer
focus and changes in developer focus as it changes over time.

@@

We now define some terms that we will use in the rest of the paper: A
\emph{message} is a block of text written by developers. In this
paper, messages will be the CVS commit comments made when the user
commits changes to files in a repository. A \emph{word distribution}
is the summary of a message by word count. Each word distribution will
be over the words in all messages. However, most words will not appear
in a message, a word distribution is effectively a word count divided
by the message size. A \emph{topic} is a word distribution. It is
essentially a set of words that form a word distribution that is
unique and independent within the set of documents in our total
corpus. One could think of a topic as a distribution of the centroid
of a group of messages. In this paper we often summarize topics by the
top 10 most frequent words of their word distribution.  A \emph{trend}
is a one or more similar topics which recur over time.  Trends are
particularly important, as they indicate long-lived and recurring
topics that may provide key insights into the development of the
project.

@@

In terms of clustering and finding topic distributions, Latent
Dirichlet Allocation (LDA)~\cite{944937} competes with Latent Semantic
Indexing
(LSI)~\cite{1421013,1374321,10.1109/ICPC.2007.13,10.1109/ICPC.2006.17},
probabilistic Latent Semantic Indexing (pLSI)~\cite{944937} and
semantic clustering~\cite{1698774,1566153}. These tools are used for
document modelling, document clustering and collaborative
filtering. LDA attempts to encode documents as mixture
model, a combination, of topics.  LDA is used in software engineering
literature~\cite{lukins2008,10.1109/MSR.2007.20,NIPS2007637,1321709}
is to extract topics from documents such as methods, bug
reports, source code and files.

@@

LSI is related to LDA and has been used to identify topics in software
artifacts for formal concept analysis and concept
location~\cite{1421013,1374321,10.1109/ICPC.2007.13,10.1109/ICPC.2006.17}.
Concept analysis aims to automatically extract concepts from source
code or documents that are related to the source code.  Concept location concerns how
high level concepts relate to low level entities such as source code. For
example, when fixing a bug how and where do the concepts in the bug
report map back to the source code?  Semantic Clustering has also been
used for similar purposes~\cite{1698774,1566153} as it is similar to
LSI.

@@

Our technique differs from the previous techniques because we apply
LDA locally to month-long windows of commit comments, where-as other
approaches apply LDA once to the entire project. This allows us to
examine the time-based windowing of topics over its development
history.

@@

In our first exploratory pass we wanted to see if LDA could provide
interesting topics extracted from a real system. We took the
repository of MySQL 3.23, extracted the commits, grouped them by 30
day non-overlapping windows, and then applied LDA to each of these
windows. We asked LDA to make 20 topics per window and we then
 examined the top 10 most frequent words in that topic.  We
found there were common words across topic clusters, such as
\emph{diffs}, \emph{annotate} and \emph{history}, these words probably should
be viewed as stop words. There were notable transitional topics such
as in the first window the word \emph{bitkeeper} appears because MySQL
adopted Bitkeeper for their source control system yet in the following
windows there were no mentions of Bitkeeper. \emph{RENAME} also only
appeared once in the first window and never again. Per each topic we
tried to identify the purpose of the topic; to our surprise we found
that even with only a little familiarity with the code base that
naming the topic was straightforward.

A sampling of the notable topic words is displayed in Table
\ref{tab:portability}, we chose topics that we felt confident we could
name.  To name each topic we selected a term that seemed to best
summarized that topic.  After extracting these topics, we attempted to
track the evolution of topics by visualizing the topics and joining
similar topics into trends.  Figure \ref{fig:lda} displays a sample
plot of the extracted topics in Table \ref{tab:portability}.

@@

2001 &  Aug &      TABLES row version \\
\hline         
2001 &  Sep &      update checksum merge \\
% 2001 &  Oct &      fixed fix \\
% 2001 &  Dec &      HPUX SCO fix \\
% \hline         
% 2002 &  Feb &      net buffer length  max\_allowed\_packet \\
% 2002 &  Mar &      small buf fix  \\
% \hline         
% 2002 &  May &      [popular] fix SCO OSF1 table\_name \\
% 2002 &  Nov &      HPUX11 compiler HP \\
% \hline         
% 2003 &  Feb &      Linux errno   \\
% 2003 &  Mar &      alarm bookmark bug \\
% \hline         
% 2003 &  Sep &      Auto logging merge windows distribution fix 64-bit 4.0 Cleanup \\
\hline
\end{tabular}
\caption{Sampled topics from MySQL 3.23, some with continuous topics. These tokens were pulled from LDA topic. Each token is a summary of one LDA generated topic from MySQL 3.23 commit logs.}
\label{tab:portability}
\end{table}

\begin{figure*}
  \centering
  \includegraphics[width=0.9\textwidth]{lda}
  \caption{Example of topics extracted from MySQL 3.23. This is the
    kind of plot we eventually want to illustrate: named topics and
    topic trends. The horizontal axis is time by month examined. The
    vertical axis is used to stack topics that occur at the same
    time. Longer topics are topics which recur in adjacent
    window. Colors are arbitrary.}
  \label{fig:lda}
\end{figure*}

\section{Methodology}
\begin{figure*}
  \centering
  \includegraphics[width=0.9\textwidth]{commit-to-topics} 
  \caption{How commits are analyzed and aggregated into Topics and Trends. Commits are first extracted, then abstracted into word counts or word distributions. These word distributions are then given a topic analysis tool like LDA. LDA finds independent word distributions (topics) that these documents are related to.}
  \label{fig:commits}
\end{figure*}

@@

\subsection{Extraction of  Repositories and Documents}

@@

@@

Given all messages, we group the messages into windows. We could use
overlapping windows, but in this paper we use non-overlapping windows
of a set time unit because it simplifies analysis.  Windowing by time
allows for many levels of granularity.  We usually used one month as
the length of our time windows. While we could use different window
lengths for the study we think that a month is a sizable unit of
development, which is large enough to show shifts in focus, but coarse
enough not to show too much detail.

@@

Once we have our data windowed, we apply our Topic Analysis tool to
each window and extract the top $N$ topics. We used $20$
topics in our case studies. Our Topic Analysis tool is an
implementation of LDA, but we could have used other similar tools like
LSI, but previous work showed more promising topic analysis results
with LDA.  We note that this step is a slow one, as executing even one
instance of LDA involves significant computation, and we perform LDA
once per window.

@@

%   6. [ ] Cluster Similarity [0/1]
\subsection{Topic Similarity}
%      1. [ ] need the tool
%             grab from the cluster analysis
%   7. [ ] Continuous Cluster Analysis

Once we have our topics extracted for each window, we analyze them and
look for topics that recur across windows. Our input for topic
similarity is the top 10 most common words of each topic.

We then cluster these topics into trends by comparing them to each
other.  Each topic is compared to each other topic in the system,
given a certain threshold of top 10 similarity, such as 8 out of 10
words. Then given this topic similarity matrix, we find the transitive
closure of similar topics; that is, if we model topics are nodes and
arcs as similarity, we fill flood along the similarity arcs until we
have partitioned the topics into clusters of similar topics. Figure
\ref{fig:closure} illustrates clustering of topics by the
topic similarity. These clusters of topics are called trends. A trend
is probably interesting if it contains more than one topic.

@@

We have devised several techniques for visualizing these topics and
trends.  For all of these techniques if we find trends that have
continuous segments, then we plot those segments as joined
horizontally across time. One technique is the \emph{compact-trend-view},
shown in Figure \ref{fig:topicsmear} and Figure \ref{fig:zoomedsmear},
which displays trends as they occur on the time-based x-axis
(placement along the y-axis is arbitrary).  Another technique is
the \emph{trend-time-line}, shown in Figure \ref{fig:trendtimeline},
each trend gets it own distinct y-value , while the x-axis is time;
these topics are then each plotted on their own line across time as
they occur. Our final technique is the \emph{trend-histogram}, shown in Figure
\ref{fig:histogram} where we plot each trend on its own line but stack
up the segments of the trend, much like a histogram.

The \emph{compact-trend-view} (Figure \ref{fig:topicsmear}) tries to
sink the larger trends to the bottom.  Once the larger trends are
stacked we fill in the gaps with the smaller trends in the same
window, and then stack the smaller trends on top.  Although there is a
chance that gaps will be left due to non-optimal stacking in practice
there are lot of small trends (90\% of all trends contain 1 topic)
which fill in these gaps quickly.  Different instances of the same
trend share the same color; apart from that, the color of trends is
randomly chosen and color similarity is not meaningful, this is easier
to see in the zoomed view of MaxDB 7.500's compact-trend-view shown in
Figure \ref{fig:zoomedsmear}.  The compact-trend-view is convenient
because it shows all of the information and the repeating continuous
trends are easy to pick out, although repeating discontinuous trends
are harder to spot.

The \emph{trend-time-line} (Figure
\ref{fig:trendtimeline}) displays repeating trends more clearly by
dedicating a horizontal line for trend segments belonging to one
trend. This means that if a trend contains discontinuous segments then
the segments appear on the same line.  However, the least common
trends need to be pruned away or the view will be very long.

The {trend-histogram} (Figure \ref{fig:histogram}) superficially
resembles the trend-time-line.  However, in this view the trends are
plotted together by stacking to the left of their row, thus time
information is lost.  The trends are ordered by the number of topics
in the trend.  The trend-histogram shows the count of instances of a
trend and thus indicates which trends occur the most. Unfortunately
due to the large number of topics given allotted space it is often
best to crop off the trends with only one topic, otherwise the tail is
long.

%XXX Describe something useful like a trend
\begin{figure*}
  \centering
  \includegraphics[width=1.0\textwidth]{fixed-time-smear-plot-scaled}
  \caption{Compact-trend-view shows topics per month for MaxDB
    7.500. The x-axis is time in months, the Y axis is used to stack
    topics occurring at the same time. Trends that are continuous are
    plotted as continuous blocks. The top 10 words in the topics are
    joined and embedded in box representing the topics.}
  \label{fig:topicsmear}
\end{figure*}

\begin{figure}
  \centering
  \includegraphics[width=0.45\textwidth]{fixed-time-smear-plot-cropped}
  \caption{A zoomed in slice of compact-trend-view of topics per month of MaxDB 7.500. The topic text is visible in each topic box. Trends are plotted across time continuously.}
  \label{fig:zoomedsmear}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=0.45\textwidth]{class-smear-plot-crop-scaled}
  \caption{Trend-time-line: Trends plotted per month of MaxDB 7.500. Time in months are plotted along the x-axis, each row on the y-axis is associated with a trend ranked by size in descending order}         
  \label{fig:trendtimeline}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=0.45\textwidth]{histogram-cropped-scaled}
  \caption{The top part of a trend-histogram of MaxDB 7.500, ordered
    by topic occurrence. X axis determines the number of continuous
    months of a trend. Trends are ranked by the number of topics that
    a trend contains in descending order.}
  \label{fig:histogram}
\end{figure}

@@

The largest common trend has references to build system files like
\texttt{SYSDD.punix} and \texttt{MONITOR.punix}.  This trend is
partially displayed at the top of the zoomed in compact-trend-view
(Figure \ref{fig:zoomedsmear}) and at the top of the trend-histogram
(Figure \ref{fig:histogram}).  Other tokens mentioned are
\emph{Sutcheck v1.09}, the prefix SUT stands for Storable Unit
Type. Sutcheck would also automate check-ins using a Perforce SCS tool
which was exporting check-ins to CVS.

@@

We noticed that Perforce and Changelist were in just about every
topic, so we added them to the stop words list. As a result, longer
trends were shortened and in some cases the total number of topics
found per month was less.  Evaluating different similarity thresholds
showed that by removing common words one reduces the general
similarity of topics. That said the larger topics still existed. Thus
if more relevant stop words are added one should tune your topic
similarity parameters to handle these changes.

@@

\subsection{Compare with topics over the entire development}

@@

We carried out the same kind of analysis, as shown in
Figure~\ref{fig:statictopics}. We extracted 20 topics  and plotted
the number of messages per month that were related to that topic. We
found was that one topic would often dominate the results, as shown , while
messages related to other topics did not appear very often.
% This technique seems useful if the 20 extracted topics are relevant
% and useful, that said one weakness of this approach is that documents
% for a topic might only occur in one time window, yet we are tracking
% messages that could be related to it over the entire lifetime of the
% project.
This approach seems reasonable if most of the extracted topics are of
broad interest during most of the development process.  However, it
may be that some topics are of strong interest but only briefly; in
such a case, a window-based model gives a much stronger indication of
the fleeting importance of such topics, and can help to put such a
topic into its proper context.

@@

If we approach the difference of static global topic analysis and
windowed topic analysis via common tokens we can see that common token
tend to dominate both results, at least in the case of MaxDB 7.500 our
local topic approach netted us topics which contained these relatively
important and common words: \emph{UNICODE}, \emph{DEC/OSF1},
\emph{ODBC}, \emph{crash}, \emph{SQLCLI}, \emph{SYSDD.cpnix},
\emph{backup}, \emph{select}, \emph{make}, \emph{memory}, \emph{view},
and finally \emph{debug}. None of these tokens were part of the static
topic analysis topics, but they were part of 566 commits (6\% of the
entire system) to MaxDB 7.500. These tokens were part of 87 out of 520
(26 months, 20 topics per month) of our locally generated topics.

@@

\begin{figure}
  \centering
  %\includegraphics[width=0.45\textwidth]{maxdb7500-everything-by-month}
  \includegraphics[width=0.45\textwidth]{month}
  \caption{MaxDB 7.500 Topics Analyzed with Static Topic Analysis, 20
    topics over the entire development history of MaxDB 7.500 (26 months). The
    shade of the topic indicates the number of documents matching that
    topic in that month relative to the number of documents (white is
    most, black is least).}
  \label{fig:statictopics}
\end{figure}

@@

We presented multiple visualizations that focused on different aspects
of the data: temporality of trends, trend size, and a
compact-trend-view. The compact-trend-view shows much more
information than the views that global analysis could show, it
indicates how focused a period is by the total number of topics as
well shows topics by similarity so one can track trends across time.
Our trend-histogram immediately hi-lights and measures the size
of trends while our trend-time-line view shows how a topic reoccurs
over time. These visualizations help us get an over feeling about the
common and unique topics that developers focus on. If implemented
properly a user can easily zoom and query a topic.
